---
title: "Fitness Data"
output: html_document
---

```{r setup, include=FALSE}
Sys.setlocale("LC_TIME","English")
knitr::opts_chunk$set(echo = TRUE)
```
##1. Loading and preprocessing the data

Unzip the data and load it in a data frame
```{r}
unzip("activity.zip")
activitydata <- read.csv("activity.csv")
```

##What is mean total number of steps taken per day?

Make a sum of all steps per day.
```{r}
library(ggplot2)
StepsPerDay <- aggregate(steps ~ date, data = activitydata, FUN = sum)
qplot(StepsPerDay$steps, main = "Steps taken per day", xlab = "steps", 
      binwidth = 500)
rmean <- mean(StepsPerDay$steps)
rmedian <- median(StepsPerDay$steps)
```

The average is `r rmean` and the median is `r rmedian`

##What is the average daily activity pattern?

Calculate the average for the steps for each interval
```{r}
stepsPerInterval<- aggregate(steps ~ interval, data = activitydata, FUN = mean)
qplot(stepsPerInterval$interval, stepsPerInterval$steps,
main = "Average daily steps", xlab = "interval",ylab = "steps", geom = "line")
selection <- max(stepsPerInterval$steps) == stepsPerInterval$steps
maxSteps <- stepsPerInterval$interval[selection]
```

The maximum steps are in this interval `r maxSteps`

## Imputing missing values

How many missing values in the data frame
```{r}
sapply(activitydata, function(x) sum(is.na(x)))
```

Only in the steps column datas are missing. This missing values are replaced,
with the steps per interval average. 

```{r}
imputed_data <- transform(activitydata, steps = ifelse(is.na(activitydata$steps), stepsPerInterval$steps[match(activitydata$interval,
                        stepsPerInterval$interval)], activitydata$steps))
imputed_dataDay <- aggregate(steps~ date, data = imputed_data, FUN = sum)
qplot(imputed_dataDay$steps, main = "Steps taken per day", xlab = "steps",
      binwidth = 500)
rimputmean <- mean(imputed_dataDay$steps)
rimputmedian <- median(imputed_dataDay$steps)
diffmean <- rimputmean - rmean
diffmedian <- rimputmedian - rmedian
```

The average is `r rimputmean` the median is `r rimputmedian`

This makes an average difference from `r diffmean`

This makes an median difference from `r diffmedian`

##Are there differences in activity patterns between weekdays and weekends?

On weekdays there is a higher peak in the morning, in the weekend there is 
more overall activity.

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
activitydata$EndOrDay <- as.factor(ifelse(is.element
                                         (weekdays(as.Date(activitydata$date))
                                                 ,weekdays)
                                         , "Weekday", "Weekend"))

stepsByInterval <- aggregate(steps ~ interval + EndOrDay, data = activitydata,
                             FUN = mean)
library(lattice)
xyplot(stepsByInterval$steps ~ stepsByInterval$interval|stepsByInterval$EndOrDay,
       main="Average Steps per Day by Interval",xlab="Interval", 
       ylab="Steps",layout=c(1,2), type="l")

```